GLS Explanation

### 2024-01-31 Notes/gls_explanation.tex
Updated simulation to incorporate single selection of theta vector.

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

@@

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Understanding Semiparametric Constraints},
  pdfauthor={Caleb Leedy},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Understanding Semiparametric Constraints}
\author{Caleb Leedy}
\date{22 January 2024}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, sharp corners]}{\end{tcolorbox}}\fi

\hypertarget{sec-intro}{%
\section{Introduction}\label{sec-intro}}

The goal of this report is to understand how additional constraints
effect semiparametric models. Inspired by Dr.~Fuller's note, I set out
to see if we could assess the amount of efficiency loss due to using
semiparametric inference. First, let's define some notation.

\hypertarget{tbl-reftab}{}
\begin{longtable}[]{@{}llll@{}}
\caption{\label{tbl-reftab}This table shows the relationship between
different quantities of interest in each segment, their estimators, and
the corresponding coefficients of their estimators.}\tabularnewline
\toprule\noalign{}
Segment & Quantity of Interest & Estimator & Coefficient \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Segment & Quantity of Interest & Estimator & Coefficient \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(A_{00}\) & \(E[g \mid X]\) & \(\hat \gamma_{11}\) & \(c_{11}\) \\
\(A_{10}\) & \(E[g \mid X]\) & \(\hat \gamma_{21}\) & \(c_{21}\) \\
\(A_{10}\) & \(E[g \mid X, Y_1]\) & \(\hat \gamma_{22}\) & \(c_{22}\) \\
\(A_{01}\) & \(E[g \mid X]\) & \(\hat \gamma_{31}\) & \(c_{31}\) \\
\(A_{01}\) & \(E[g \mid X, Y_2]\) & \(\hat \gamma_{33}\) & \(c_{33}\) \\
\(A_{11}\) & \(E[g \mid X]\) & \(\hat \gamma_{41}\) & \(c_{41}\) \\
\(A_{11}\) & \(E[g \mid X, Y_1]\) & \(\hat \gamma_{42}\) & \(c_{42}\) \\
\(A_{11}\) & \(E[g \mid X, Y_2]\) & \(\hat \gamma_{43}\) & \(c_{43}\) \\
\(A_{11}\) & \(E[g \mid X, Y_1, Y_2]\) & \(\hat \gamma_{44}\) &
\(c_{44}\) \\
\end{longtable}

This is a similar framework to understanding the class of estimators
that we discussed in \verb|optest_update.pdf|, where we write the
estimator as a weighted average of different expectations with
functional coefficients.

\[\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta)E[g(Z)
\mid X] + \beta_1(\delta)E[g(Z) \mid X, Y_1] + \beta_2(\delta) E[g(Z)
\mid X, Y_2].\]

In this case, we assume that \(E[\beta_j(\delta)] = 0\) for
\(j = 0, 1, 2\).

However, now we write the estimator as

\begin{equation}\protect\hypertarget{eq-linest}{}{\hat \theta = \sum_{k,t} c_{kt}\hat \gamma_{kt}}\label{eq-linest}\end{equation}

where
\(\hat \gamma_{kt} = \frac{\delta_{ij}}{\pi_{ij}}E[g \mid G_{ij}(X, Y_1, Y_2)]\)
and \(ij\) corresponds to the segment \(A_{ij}\) associated with
\(\hat \gamma_{kt}\) in Table~\ref{tbl-reftab}. In general the estimator
has the following form,

\begin{align*}
\hat \theta = n^{-1} \sum_{i = 1}^n 
  &\left\{c_{11}\frac{\delta_{00}}{\pi_{00}} E[g \mid X]+ \right.\\
  &c_{21}\frac{\delta_{10}}{\pi_{10}} E[g \mid X]+ 
  c_{22}\frac{\delta_{10}}{\pi_{10}} E[g \mid X, Y_1]+ \\
  &c_{31}\frac{\delta_{01}}{\pi_{01}} E[g \mid X]+ 
  c_{33}\frac{\delta_{01}}{\pi_{01}} E[g \mid X, Y_2]+ \\
  &c_{41}\frac{\delta_{11}}{\pi_{11}} E[g \mid X]+ 
  c_{42}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_1]+
  c_{43}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_2]+
  \left.c_{44}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_1, Y_2]\right\}
\end{align*}

which we have condensed to Equation~\ref{eq-linest}.

\hypertarget{sec-methods}{%
\section{Methods}\label{sec-methods}}

As noted by Dr.~Fuller, we can understand parametric estimation as
solving the following constrained optimization problems:

\begin{align*}
  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
  \sum_{k,t} c_{kt} = 1 \\
       &\text{or} \\
  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
  \sum_{(k,t): (k,t) \neq (4, 4)} c_{kt} = 0 \text{ and } c_{44} = 1. \\
\end{align*}

To be robust to the outcome model, we can construct a similar
optimization problem with different constraints:

\begin{align*}
  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
  c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0,\\
       &\text{and } c_{44} = 1.
\end{align*}

Likewise to be robust to the response model, we can have the problem:

\begin{align*}
  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
  c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \\
       &\text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}.
\end{align*}

To be double robust (robust to the outcome and response model) we can
combine the last two constraints. This is summarized in
Table~\ref{tbl-mods}.

\hypertarget{tbl-mods}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2941}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7059}}@{}}
\caption{\label{tbl-mods}This table identifies the different constraints
for each model type.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Type
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Constraints
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Type
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Constraints
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Parametric 1 & \(\sum_{k, t} c_{kt} = 1\) \\
Parametric 2 &
\(\sum_{k, t: (k, t) \neq (4, 4)} c_{kt} = 0, c_{44} = 1\) \\
Outcome Robust &
\(c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, \text{ and } c_{44} = 1.\) \\
Response Robust &
\(c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}\) \\
Double Robust &
\(c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, c_{44} = 1, c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}\) \\
\end{longtable}

We can justify the constraints for the Outcome Robust model by rewriting
the estimator into the following form:

\begin{align*}
E_\delta[\hat \theta] &= n^{-1} \sum_{i = 1}^n E_\delta\left[
E[g \mid X] \left(\frac{\delta_{00}}{\pi_{00}} c_{11} +
\frac{\delta_{10}}{\pi_{10}}c_{21} \frac{\delta_{01}}{\pi_{01}}c_{31} +
\frac{\delta_{11}}{\pi_{11}}c_{41}\right)\right. + \\ 
&\qquad\qquad\qquad\qquad E[g \mid X, Y_1] \left(\frac{\delta_{10}}{\pi_{10}}c_{22} + 
\frac{\delta_{11}}{\pi_{11}}c_{42}\right) +\\ 
&\qquad\qquad\qquad\qquad E[g \mid X, Y_2] \left(\frac{\delta_{01}}{\pi_{01}}c_{33} + 
\frac{\delta_{11}}{\pi_{11}}c_{43}\right) +\\ 
&\qquad\qquad\qquad\qquad \left.g \frac{\delta_{11}}{\pi_{11}}c_{44}\right] \\
&= n^{-1} \sum_{i = 1}^n \left\{E[g \mid X](c_{11} + c_{21} + c_{31} + c_{41}) +
E[g \mid X, Y_1] (c_{22} + c_{42}) +
E[g \mid X, Y_2] (c_{33} + c_{43}) + g c_{44}\right\}
\end{align*}

Thus, if the Outcome Robust constraints are satisfied,
\(E_\delta[\hat \theta] = g\) when we take the expectation with respect
to the response variable \(\delta\). These constraints make sense
because if they are satisfied \(E[\hat \theta]\) does not depend on the
correct specification of the models for \(E[g \mid G_r(X, Y_1, Y_2)]\).

Likewise, we can justify the constraints for the Response Robust model
by rewriting Equation~\ref{eq-linest} into the form

\begin{align*}
E_Z [\hat \theta] &= n^{-1} \sum_{i = 1}^n E_Z\left[
\frac{\delta_{00}}{\pi_{00}}c_{11} E[g \mid X]\right. + \\
&\qquad\qquad\qquad\qquad\frac{\delta_{10}}{\pi_{10}}\left(c_{21}E[g \mid X] + 
  c_{22}E[g \mid X, Y_1]\right) +\\
&\qquad\qquad\qquad\qquad\frac{\delta_{01}}{\pi_{01}}\left(c_{31}E[g \mid X] + 
  c_{33}E[g \mid X, Y_2]\right) +\\
&\qquad\qquad\qquad\qquad\left.\frac{\delta_{11}}{\pi_{11}}\left(c_{41}E[g \mid X] + 
  c_{42}E[g \mid X, Y_1] + c_{43}E[g \mid X, Y_2] + c_{44}g\right)\right] \\
& n^{-1} \sum_{i = 1}^n \left(\frac{\delta_{00}}{\pi_{00}}c_{11} + 
\frac{\delta_{10}}{\pi_{10}}(c_{21} + c_{22}) + 
\frac{\delta_{01}}{\pi_{01}}(c_{31} + c_{33}) + 
\frac{\delta_{11}}{\pi_{11}}(c_{41} + c_{42} + c_{43} + c_{44})\right)g
\end{align*}

Thus, if the Response Robust constraints are satisfied,
\(E_X[\hat \theta] = g\) when we take the expectation with respect to
the outcome variable \(Z = (X, Y_1, Y_2)'\). These constraints make
sense because when they are satisfied \(E[\hat \theta]\) does not depend
on the correct response probabilities \(\pi = E[\delta]\).

\hypertarget{sec-simulations}{%
\section{Simulation Study}\label{sec-simulations}}

\hypertarget{simulation-1}{%
\subsection*{Simulation 1}\label{simulation-1}}
\addcontentsline{toc}{subsection}{Simulation 1}

We use the following simulation setup

\begin{align*}
  \begin{bmatrix} x \\ e_1 \\ e_2 \end{bmatrix} 
  &\stackrel{ind}{\sim} N\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
  \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & \rho \\ 0 & \rho & 1 \end{bmatrix}\right) \\
  y_1 &= x + e_1 \\
  y_2 &= \theta + x + e_2 \\
\end{align*}

Furthermore, \(\pi_{11} = 0.2\) and \(\pi_{00} = 0.3, \pi_{10} = 0.4\),
and \(\pi_{01} = 0.1\). The goal of this simulation study is to find
\(\theta = E[Y_2]\). In other words, \(g(Z) = Y_2\). There are several
algorithms for comparison which are defined as the following:

\begin{align*}
  Oracle &= n^{-1} \sum_{i = 1}^n g(Z_i)\\
  CC &= \frac{\sum_{i = 1}^n \delta_{11} g(Z_i)}{\sum_{i = 1}^n \delta_{11}} \\
  IPW &= \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g(Z_i)\\
\end{align*}

Furthermore, we include four existing estimators that we have already
proposed in the past: WLS, Prop, PropInd, and SemiDelta.

The estimator WLS is a a weight least square estimator derived in the
following manner. We now consider a normal model:

\[ 
  \begin{pmatrix}
    x_i \\ e_{1i} \\ e_{2i}
  \end{pmatrix} \stackrel{ind}{\sim}
  N \left(
  \begin{bmatrix}
    0 \\ 0 \\ 0
  \end{bmatrix},
  \begin{bmatrix}
    1 & 0 & 0 \\
    0 & \sigma_{11} & \sigma_{12} \\ 
    0 & \sigma_{12} & \sigma_{22}
  \end{bmatrix},
  \right)
\]

and define \(y_{1i} = \theta_1 + x_i + e_{1i}\) and
\(y_{2i} = \theta_2 + x_i + e_{2i}\). Then \(b_1 = b_2 = 1\). We define
\(\bar z_k^{(ij)}\) as the mean of \(y_k\) in segment \(A_{ij}\). This
means that we have means \(\bar z_1^{(11)}\), \(\bar z_2^{(11)}\),
\(\bar z_1^{(10)}\), and \(\bar z_2^{(01)}\). Let
\(W = [\bar z_1^{(11)}, \bar z_2^{(11)}, \bar z_1^{(10)}, \bar z_2^{(01)}]'\),
then for \(n_{ij} = |A_{ij}|\), we have

\[Z - M \mu \sim N(\vec 0, V)\]

where

\[M = 
  \begin{bmatrix}
    1 & 0 \\
    0 & 1 \\
    1 & 0 \\
    0 & 1 \\
  \end{bmatrix}
  \text{ and }
  V = 
  \begin{bmatrix}
    \frac{\sigma_{11}}{n_{11}} & \frac{\sigma_{12}}{n_{11}} & 0 & 0 \\
    \frac{\sigma_{12}}{n_{11}} & \frac{\sigma_{22}}{n_{11}} & 0 & 0 \\
    0 & 0 & \frac{\sigma_{11}}{n_{10}} & 0 \\
    0 & 0 & 0 & \frac{\sigma_{22}}{n_{01}} \\
  \end{bmatrix}.
\]

Thus, the BLUE for \(\mu = [\mu_1, \mu_2]'\) is

\[\hat \mu = (M' V^{-1} M)^{-1} M' V^{-1} W.\]

Hence, WLS is \(\mu_2\) as \(g(X, Y_1, Y_2) = Y_2\).

The remaining three estimators are drived from the following expression
where the values for \(\beta\) are provided in Table~\ref{tbl-beta}.
These are good estimators to compare to because Prop is the original
proposed estimator. PropInd has the same form as Prop except the values
for \(\beta\) is different. PropInd shares the same values of \(\beta\)
as all of the new models with constraints. SemiDelta is useful because
it is the best estimator in general so far.

\[
\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + 
\beta_0(\delta, c_0)E[g(Z) \mid X] + 
\beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + 
\beta_2(\delta, c_2) E[g(Z)
\mid X, Y_2].
\]

\hypertarget{tbl-beta}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1698}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4151}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4151}}@{}}
\caption{\label{tbl-beta}This table displays the values of \(\beta\) for
different estimator types.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Estimator
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(\beta_0(\delta, c_0)\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(\beta_1(\delta, c_1)\)
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Estimator
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(\beta_0(\delta, c_0)\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(\beta_1(\delta, c_1)\)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Prop &
\(\left(1 - \frac{(\delta_{10} + \delta_{11})}{(\pi_{10} + \pi_{11})} - \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)\)
&
\(\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)\) \\
PropInd &
\(\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - \frac{(\delta_{01})}{(\pi_{01})} + \frac{\delta_{11}}{\pi_{11}}\right)\)
&
\(\left(\frac{\delta_{10}}{\pi_{10}} - \frac{\delta_{11}}{\pi_{11}}\right)\) \\
SemiDelta &
\(c_0\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{00}}{\pi_{00}}\right)\)
&
\(c_1\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{10}}{\pi_{10}}\right)\) \\
\end{longtable}

In the simulation results in Table~\ref{tbl-sim1}, the new results have
the same label as the value from the \texttt{Type} column in
Table~\ref{tbl-mods}.

\hypertarget{tbl-sim1}{}
\begin{longtable}[]{@{}lrrrr@{}}
\caption{\label{tbl-sim1}Results from Simulation 1. True values:
\(\theta = 5, \rho = 0.5\). The test conducted for the T-statistic and
P-value is a two sample test to see if the estimator is unbiased. This
simulation uses \(Y_1 = x + \varepsilon_1\),
\(Y_2 = \theta + x + \varepsilon_2\) where \(X \sim N(0, 1)\) and
\((\varepsilon_1, \varepsilon_2)\) come from a mean zero bivariate
normal distribution with unit variance and covariance \(\rho\). The
segments are unbalanced with: \(\pi_{11} = 0.2\), \(\pi_{10} = 0.4\),
\(\pi_{01} = 0.1\), and \(\pi_{00} = 0.3\).}\tabularnewline
\toprule\noalign{}
Algorithm & bias & sd & tstat & pval \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Algorithm & bias & sd & tstat & pval \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Oracle & -0.002 & 0.045 & -1.669 & 0.048 \\
CC & 0.001 & 0.083 & 0.263 & 0.396 \\
IPW & 0.010 & 0.357 & 0.872 & 0.192 \\
WLS & 0.000 & 0.054 & -0.065 & 0.474 \\
\(\hat \theta_{prop}\) & -0.002 & 0.063 & -0.768 & 0.221 \\
\(\hat \theta_{\delta}\) & -0.001 & 0.064 & -0.342 & 0.366 \\
Parametric 1 & -0.001 & 0.061 & -0.509 & 0.305 \\
Parametric 2 & -0.001 & 0.061 & -0.509 & 0.305 \\
Outcome Robust & -0.001 & 0.061 & -0.543 & 0.294 \\
Response Robust & -0.001 & 0.061 & -0.509 & 0.305 \\
Double Robust & -0.001 & 0.061 & -0.543 & 0.294 \\
\end{longtable}

\hypertarget{simulation-2}{%
\subsection*{Simulation 2}\label{simulation-2}}
\addcontentsline{toc}{subsection}{Simulation 2}

Next, we used the same simulation setup except that we are interested in
estimating \(\theta = E[g] = E[Y_1^2 Y_2]\). We do not use the WLS
estimator because it does not work.

>>>>

@@

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}
\renewcommand{\arraystretch}{1.4}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={GLS Explanation},
  pdfauthor={Caleb Leedy},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{GLS Explanation}
\author{Caleb Leedy}
\date{31 January 2024}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, breakable]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

This is a write up explaning the dual form of generalized least squares
(GLS). We think of GLS as an extension of ordinary least squares (OLS),
so I will discuss this first and then comment on how these are connected
and the dual formulation for both.

\hypertarget{the-gauss-markov-theorem}{%
\section{The Gauss-Markov Theorem}\label{the-gauss-markov-theorem}}

It all comes back to the Gauss-Markov Theorem, which states that for a
linear model in the form,

\[ Y = X\beta + \varepsilon\]

with \(E[\varepsilon] = 0\) and \(\Var(\varepsilon) = \sigma^2 I_n\),
the BLUE (best linear unbiased estimator) of \(C\beta\) for \(C = AX\)
for some matrix \(A\) is (assuming \(X'X\) has an inverse),

\[ C\hat \beta = C(X'X)^{-1} X'Y = AX(X'X)^{-1}X'Y.\]

\hypertarget{gls-extension}{%
\subsection{GLS Extension}\label{gls-extension}}

To extend this to the GLS setting consider the case where
\(\Var(\varepsilon) = \sigma^2 V\) for some known positive-definite
matrix \(V\). In this case, we can rewrite \(Y = X\beta + \varepsilon\)
as

\[ Z = W\beta + \tilde \varepsilon\]

where \(Z = V^{-1/2} Y\), \(W = V^{-1/2}X\) and
\(\tilde \varepsilon = V^{-1/2} \varepsilon\). But this formulation
implies that \(E[\tilde \varepsilon] = 0\) and
\(\Var(\tilde \varepsilon) = \sigma^2 I_n\) and so hence by the
Gauss-Markov Theorem, the BLUE is

\[ (W'W)^{-1} W'Z = (X'V^{-1}V)^{-1} X'V^{-1}Y.\]

\hypertarget{dual-form}{%
\section{Dual Form}\label{dual-form}}

All of this comes back to the fact that we want to estimate the BLUE.
However, we could do this directly. By definition, the BLUE is the
minimum variance linear unbiased estimator. This means that we consider
a class of estimators \(CW'Y\) such that \(E[CW'Y] = C\beta\) that
minimizes the variance \(CW'\Var(Y) WC' \propto CW'WC'\). Since we
assume the model form of \(Y = X\beta + \varepsilon\), the unbiasedness
condition is equivalent to,

\[ E[CW'Y] = CW'E[Y] = CW'X\beta = C\beta. \]

Since this is true for all \(\beta\), it must be the case that

\[ W'X = I_p \text{ for } p = \dim(\beta).\]

Hence, we have a Lagrangian

\[L(W) \propto CW'WC' - (W'X - I_p)\lambda.\]

Thus, the first-order conditions are \(W - X\lambda = 0\) and
\(W'X - I_p = 0\). This yields a solution of \(W = X(X'X)^{-1}\) which
is exactly what we get from OLS because the final result is
\(CW'Y = C(X'X)^{-1} X'Y\). To get the GLS form, we can keep the same
constraint except minimize the variance \(W'VW\).

\hypertarget{estimating-the-variance}{%
\section{Estimating the Variance}\label{estimating-the-variance}}

When estimating GLS, we assume that the covariance matrix \(V\) is fully
known. What do we do if we have to estimate \(V\)? In this case, we need
to use (finite-dimensional) semiparametric theory to estimate the
optimal \(W\) such that \(W\) belongs to the orthogonal nuisance tangent
space of \(\hat V\). This is something that we could explore further if
we want to.
